The Concurrent Calculi Formalisation Benchmark

### 2024-02-23 challenges/coinduction.tex
Use title case for LNCS guideline compliance.

\subsection{Challenge: Coinduction and infinite processes}
% and reasoning about infinite processes}
\label{sec:challenge:coinduction}

@@

\subsubsection{Bisimilarity.}
Bisimilarity is a notion of equivalence for processes and
% .  Our notion of process equivalence relations
builds on a notion of \emph{observables}, i.e., what we can externally
observe from the semantics of a process. If we allowed ourselves only to observe
internal transitions (i.e., observe that a process is internally
performing a step of computation) we would relate either too few
processes (in the strong case where we relate only processes with
exactly the same number of internal transitions) or every process (in
the weak case where we relate processes with any amount of internal
transitions).  As a result, we must allow ourselves to observe more than
just internal transitions, and we choose to describe a process's observables
as the names it might use for sending and receiving.

@@

\subsubsection{Strong barbed congruence.}
In order to detect cases like the one above, we need to restrict strong barbed bisimilarity so that it becomes a congruence,
i.e., % we need to make strong barbed bisimilarity consider
we have to consider the environment in which processes may be placed.
% (i.e.\ the context) of processes.

>>>>

\subsection{Challenge: Coinduction and Infinite Processes}
% and reasoning about infinite processes}
\label{sec:challenge:coinduction}

@@

\subsubsection{Strong Barbed Bisimilarity.}
Bisimilarity is a notion of equivalence for processes and
% .  Our notion of process equivalence relations
builds on a notion of \emph{observables}, i.e., what we can externally
observe from the semantics of a process. If we allowed ourselves only to observe
internal transitions (i.e., observe that a process is internally
performing a step of computation) we would relate either too few
processes (in the strong case where we relate only processes with
exactly the same number of internal transitions) or every process (in
the weak case where we relate processes with any amount of internal
transitions).  As a result, we must allow ourselves to observe more than
just internal transitions, and we choose to describe a process's observables
as the names it might use for sending and receiving.

@@

\subsubsection{Strong Barbed Congruence.}
In order to detect cases like the one above, we need to restrict strong barbed bisimilarity so that it becomes a congruence,
i.e., % we need to make strong barbed bisimilarity consider
we have to consider the environment in which processes may be placed.
% (i.e.\ the context) of processes.

### 2024-02-23 challenges/linearity.tex
Use title case for LNCS guideline compliance.

\subsection{Challenge: Linearity and behavioural type systems}
\label{sec:challenge:linearity-beh-types}
This challenge formalises a proof that requires reasoning about the linearity of channels.
Linearity is the notion that a channel must be used exactly once by a process.
This is necessary to prove properties about session type systems, and the key
issue of this challenge is reasoning about the linearity of context splitting operations.
Linear reasoning is also necessary to formalise, \eg, linear and affine types for the \(\pi\)-calculus and cut elimination in linear logics.

@@

\subsubsection{Session types.}
Our process syntax allows us to write processes that are ill formed in
the sense that they either use the endpoints bound by a restriction to
communicate in a way that does not follow the intended duality, or
attempt to send something which is not a base value.  As an example,
the process
\( \Presd{x}{y}{(\Ppar{\Pout{x}{a}{\Pend}}{\Pout{y}{a}{\Pend}})} \)
attempts to send a base value on both \( x \) and \( y\), whereas one
of the names should be used for receiving in order to guarantee
progress.  Another example is the process
\( \Presd{x}{y}{(\Ppar{\Pout{x}{l}{\Pend}}{\Pin{y}{l}{\Pend}})} \),
which attempts to send a variable that is not instantiated at the time
of sending.
%
To prevent these issues, we introduce a \emph{session type system} which
rules out ill-formed processes.

@@

\subsubsection{Typing rules.}
Our type system is aimed at maintaining two invariants:
\begin{enumerate}
\item No endpoint is used simultaneously by parallel processes;
\item The two endpoints of the same session have dual types.
\end{enumerate}
The first invariant is maintained by linearly splitting type contexts when typing compositions of processes, the second by requiring duality when typing restrictions.

>>>>

\subsection{Challenge: Linearity and Behavioural Type Systems}
\label{sec:challenge:linearity-beh-types}
This challenge formalises a proof that requires reasoning about the linearity of channels.
Linearity is the notion that a channel must be used exactly once by a process.
This is necessary to prove properties about session type systems, and the key
issue of this challenge is reasoning about the linearity of context splitting operations.
Linear reasoning is also necessary to formalise, \eg, linear and affine types for the \(\pi\)-calculus and cut elimination in linear logics.

@@

\subsubsection{Session Types.}
Our process syntax allows us to write processes that are ill formed in
the sense that they either use the endpoints bound by a restriction to
communicate in a way that does not follow the intended duality, or
attempt to send something which is not a base value.  As an example,
the process
\( \Presd{x}{y}{(\Ppar{\Pout{x}{a}{\Pend}}{\Pout{y}{a}{\Pend}})} \)
attempts to send a base value on both \( x \) and \( y\), whereas one
of the names should be used for receiving in order to guarantee
progress.  Another example is the process
\( \Presd{x}{y}{(\Ppar{\Pout{x}{l}{\Pend}}{\Pin{y}{l}{\Pend}})} \),
which attempts to send a variable that is not instantiated at the time
of sending.
%
To prevent these issues, we introduce a \emph{session type system} which
rules out ill-formed processes.

@@

\subsubsection{Typing Rules.}
Our type system is aimed at maintaining two invariants:
\begin{enumerate}
\item No endpoint is used simultaneously by parallel processes;
\item The two endpoints of the same session have dual types.
\end{enumerate}
The first invariant is maintained by linearly splitting type contexts when typing compositions of processes, the second by requiring duality when typing restrictions.

### 2024-02-23 challenges/scope-extrusion.tex
Use title case for LNCS guideline compliance.

\subsection{Challenge: Name passing and scope extrusion}
\label{sec:challenge:name-passing-scope-extrusion}

@@

\subsubsection{Reduction semantics.}
The first semantics is an operational reduction semantics, which avoids reasoning explicitly about scope extrusion by way of a structural congruence rule.
\emph{Structural congruence} is the smallest congruence relation that satisfies the following axioms:

@@

\subsubsection{Transition system semantics.}
The second semantics of the system describe the actions that the system can perform by defining a labelled transition relation on processes.
The transitions are labelled by \emph{actions}, the syntax of which is:
\[
\begin{array}{r@{\qquad}c@{\qquad}l}
  \alpha & \Coloneqq & \Aoutf{x}{y} \quad\mid\quad \Ain{x}{y} \quad\mid\quad \Aoutb{x}{y} \quad\mid\quad \Atau
\end{array}
\]

>>>>

\subsection{Challenge: Name Passing and Scope Extrusion}
\label{sec:challenge:name-passing-scope-extrusion}

@@

\subsubsection{Reduction Semantics.}
The first semantics is an operational reduction semantics, which avoids reasoning explicitly about scope extrusion by way of a structural congruence rule.
\emph{Structural congruence} is the smallest congruence relation that satisfies the following axioms:

@@

\subsubsection{Transition System Semantics.}
The second semantics of the system describe the actions that the system can perform by defining a labelled transition relation on processes.
The transitions are labelled by \emph{actions}, the syntax of which is:
\[
\begin{array}{r@{\qquad}c@{\qquad}l}
  \alpha & \Coloneqq & \Aoutf{x}{y} \quad\mid\quad \Ain{x}{y} \quad\mid\quad \Aoutb{x}{y} \quad\mid\quad \Atau
\end{array}
\]

### 2024-02-23 short-paper/short-paper.tex
Use title case for LNCS guideline compliance.

\section{Overview and design considerations}\label{sec:design-discussion}
In this section, we outline the factors considered when designing the benchmark challenges.
We begin with some general remarks, then describe the individual design considerations for each challenge problem, and the criteria for evaluating solutions thereof.

@@

\subsection{Scope extrusion}
This challenge revolves around the mechanisation of scope
extrusion, by which a process can send restricted names to another
process, as long as the restriction can safely be extruded to include
the receiving process.  The setting for this challenge is a
``classic'' untyped \( \pi \)-calculus, where (unlike the calculi in
the other challenges) names can be sent and received, and bound by
input constructs.  We define two different semantics for our system:
\begin{enumerate}
\item A reduction system: this avoids explicit reasoning about scope
  extrusion by using instead structural congruence, allowing
  implementors to explore alternative ways to encode the latter (\eg
  via process contexts or via compatible refinement);
\item An (early) labelled transition system.
\end{enumerate}
The objective of this
challenge is to prove that the two semantics are equivalent up to
structural congruence.

@@

\subsection{Evaluation criteria.}
The motivation behind our benchmark is to obtain evidence towards answering
questions \cref{item:rq1,item:rq2,item:rq3}. We are therefore
interested not only in the solutions, but also in the experience of
solving the challenges with the chosen approach.  Solutions to our
challenges should be compared on three axes:
\begin{enumerate}
\item Mechanisation overhead: the amount of manually-written infrastructure and setup needed to express the definitions in the mechanisation;
\item Adequacy of the formal statements in the mechanisation: whether the proven theorems are easily recognisable as the theorems from the challenge; and
\item Cost of entry for the tools and techniques employed: the difficulty of learning to use the techniques.
\end{enumerate}
Solutions to our challenges need not strictly follow the definitions and lemmata set out in the problem text, but solutions which deviate from the original challenges must present more elaborate argumentation for their adequacy.

@@

% We previously promise to discuss combinations of techniques and applications to more elaborate schemes
%\section{Going Beyond the Challenge Problems}\label{sec:going-beyond}
\section{Future work and conclusions}\label{sec:going-beyond}
Our benchmark challenges do not cover all issues in the field, but focus on the fundamental aspects of linearity, scope extrusion, and coinduction.
Many mechanisations need to combine techniques to handle several of these aspects at once, and some may also need to handle aspects that are not covered by our benchmark challenges at all.

>>>>

\section{Overview and Design Considerations}\label{sec:design-discussion}
In this section, we outline the factors considered when designing the benchmark challenges.
We begin with some general remarks, then describe the individual design considerations for each challenge problem, and the criteria for evaluating solutions thereof.

@@

\subsection{Scope Extrusion}
This challenge revolves around the mechanisation of scope
extrusion, by which a process can send restricted names to another
process, as long as the restriction can safely be extruded to include
the receiving process.  The setting for this challenge is a
``classic'' untyped \( \pi \)-calculus, where (unlike the calculi in
the other challenges) names can be sent and received, and bound by
input constructs.  We define two different semantics for our system:
\begin{enumerate}
\item A reduction system: this avoids explicit reasoning about scope
  extrusion by using instead structural congruence, allowing
  implementors to explore alternative ways to encode the latter (\eg
  via process contexts or via compatible refinement);
\item An (early) labelled transition system.
\end{enumerate}
The objective of this
challenge is to prove that the two semantics are equivalent up to
structural congruence.

@@

\subsection{Evaluation Criteria.}
The motivation behind our benchmark is to obtain evidence towards answering
questions \cref{item:rq1,item:rq2,item:rq3}. We are therefore
interested not only in the solutions, but also in the experience of
solving the challenges with the chosen approach.  Solutions to our
challenges should be compared on three axes:
\begin{enumerate}
\item Mechanisation overhead: the amount of manually-written infrastructure and setup needed to express the definitions in the mechanisation;
\item Adequacy of the formal statements in the mechanisation: whether the proven theorems are easily recognisable as the theorems from the challenge; and
\item Cost of entry for the tools and techniques employed: the difficulty of learning to use the techniques.
\end{enumerate}
Solutions to our challenges need not strictly follow the definitions and lemmata set out in the problem text, but solutions which deviate from the original challenges must present more elaborate argumentation for their adequacy.

@@

% We previously promise to discuss combinations of techniques and applications to more elaborate schemes
%\section{Going Beyond the Challenge Problems}\label{sec:going-beyond}
\section{Future Work and Conclusions}\label{sec:going-beyond}
Our benchmark challenges do not cover all issues in the field, but focus on the fundamental aspects of linearity, scope extrusion, and coinduction.
Many mechanisations need to combine techniques to handle several of these aspects at once, and some may also need to handle aspects that are not covered by our benchmark challenges at all.
